A family of accessibility measures derived from spatial interaction principles

### 2025-05-23 paper/manuscript.tex
Cleaned up a few small equation errors - and posted preprint link to repo page.

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}
\usepackage{draftwatermark}
\SetWatermarkText{DRAFT}
\SetWatermarkScale{1.5}
\SetWatermarkColor[gray]{0.95}
\usepackage{float}
\floatplacement{table}{H}
\floatplacement{figure}{H}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A family of accessibility measures derived from spatial interaction principles},
  pdfauthor={Anastasia Soukhov*; Rafael H. M. Pereira; Christopher D. Higgins; Antonio Páez},
  pdfkeywords={accessibility; spatial interaction model; balancing
factors; competitive accessibility; constraints; singly constrained;
doubly constrained; total constrained},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{A family of accessibility measures derived from spatial
interaction principles}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Submitted preprint for peer-review (May 2025)}
\author{Anastasia Soukhov* \and Rafael H. M. Pereira \and Christopher D.
Higgins \and Antonio Páez}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Transportation planning has long prioritized the efficiency of movement,
or mobility. However, the concept of accessibility represents a more
comprehensive evolution, shifting focus from mere movement to the
potential to reach (i.e., spatially interact) with desired destinations.
Despite growing recognition of accessibility-based planning approaches,
the concept remains fragmented, with inconsistent definitions and
unclear interpretations. This work's aim is to clarify and unify the
concept of accessibility by connecting it into spatial interaction
modeling. We demonstrate that widely used mobility and accessibility
models, such as gravity-based accessibility and spatial interaction
models, share common theoretical roots. From this foundation, this paper
offers three contributions: (A) we introduce a family of accessibility
measures within the principles of spatial interaction, and (B) formally
define four members of the family, namely the `unconstrained' measure
(i.e., Hansen-type accessibility), the `total constrained' measure
(i.e., a constrained version of the Hansen-type accessibility), the
`singly constrained' measure (i.e., related to the popular two step
floating catchment approach - 2SFCA), and the `doubly constrained'
measure representing realized interactions or `access', effectively
equal to the doubly constrained spatial interaction model; and (C) we
demonstrate the interpretability advantages of the family, as these
constrained accessibility measures yield values in units of the number
of potential ``opportunities for spatial interaction'' or ``population
for spatial interaction'' for each zone and zonal flow. The family of
accessibility measures proposed here clarifies the concept of
`potential' in accessibility, demonstrates theoretical and formulaic
linkages across popular accessibility and spatial interaction models,
and reintroduces measurement units into accessibility measures. By doing
so, we believe this family of measures can unlock a clearer, more
interpretable, and cohesive foundation for accessibility analysis.

*Email of corresponding author for preprint: soukhoa@mcmaster.ca
\end{abstract}

@@

Understanding \(S_i\) as the inverse of Wilson's \(A_i\) does not
uncover any new meaning for \(S_i\) itself. Indeed, mathematically it is
true, \(A_i\)'s role in Wilson's general model
\(T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})\) is that of a balancing
factor \(k\) i.e., keeping units balanced and proportionality based on
constraints. Understanding \(S_i\) itself as a balancing factor is not
wholly helpful as Hansen and (Stewart before him) defined accessibility
as a partial sum of the demographic force \(F\) or the system-wide
population potential of opportunities for spatial interaction (i.e.,
(\textbf{q-stewart-population-potential-sum?}) representing
\(V_i = \sum_j \frac{M_j}{d_{ij}}\) after losing \(G\)).

@@

Firstly, Table~\ref{tbl-small-system-land-use} summarises the population
(in units of 10,000s of people) and the opportunities (the number of
physicians) per zone. {[}\^{}1{]}. Considering
Table~\ref{tbl-small-system-land-use}'s values, the
Provider-to-Population-Ratio (PPR) in this system is 24.5. For
reference, the number of physicians per 10,000 in Canada in 2022 was
24.97 (WHO 2025). Secondly, to pair the zonal population and opportunity
information, the assumed cost of movement (in minutes of travel time)
between origins and destinations is as shown in
Table~\ref{tbl-small-system-cost}.

@@

Further, we can see that, since \(D\) and \(W^{(2)}_j\) are both in
units of opportunities, the proportional allocation factor for the total
constrained opportunity case \(\kappa_i^T\) is dimensionless:

\[
\kappa_i^T = \frac{\sum_j D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

\noindent and therefore \(V^T_i\) is now in the units of \(W^{(2)}_j\),
that is, the mass at the destination
(\(V^T_i = \kappa_i^T\sum_j W^{(2)}_j\)). The role of \(\kappa_i^T\) in
this reformulation of accessibility is to transform between units and to
adjust the number of opportunities accessible from \(i\) so they
represent a proportion of the total number of opportunities in the
region. \(\kappa_i^T\) then assigns opportunities in proportion to the
impedance between \(i\) and \(j\). This is why we refer to
\(\kappa_i^T\) as a proportional allocation factor. On the other hand,
the proportionality constant \(K^T\) balances the units of \(V^0_i\),
the Hansen-type accessibility values, and is an alternative expression
of the total constrained accessibility measure.

@@

Using the calculated proportionality constants for all zones and
multiplying them by the unconstrained accessibility value \(V^0_i\), the
total opportunity constrained accessibility values for all zones and
different travel behaviour scenarios is presented in
Table~\ref{tbl-simple-example-total-opportunity-accessibility}.
\(\kappa_i^T\) for each zone is not shown, but can be understood to be
the unitless proportion of opportunities (of the total opportunities)
allocated to each zone based on travel impedance.

@@

\caption{\label{tbl-simple-example-total-opportunity-accessibility}Simple
system: total opportunity constrained accessibility.}

@@

In contrast to unconstrained accessibility \(V^0_i\), imposing a
constraint -in this case a total opportunity constraint for this
variant- allows for the comparison of differences and ratios between
regions and across different travel behaviour scenarios as well. Each
value is effectively in units of physicians, with the impedance units
already accounted for by \(\kappa_i^T\).

@@

One can also directly compare values at a specific zone due to the
consistent units. For instance, zone 1 remains intermediate in capturing
accessible physicians relative to zones 2 and 3 across scenarios,
similar to the unconstrained case. However, the difference between
travel behaviour scenarios differ in direction. Specifically, Zone 1
captures 0.6067946 more and 7.9850478 fewer opportunities than the
lesser decay scenarios \(f_2(c_{ij})\) and \(f_3(c_{ij})\) respectively.
Why? \(\kappa_i^T\) ensures proportional allocation for each travel
behaviour scenario. Meaning, while the unconstrained accessibility
increases, \(\kappa_i^T\) adjusts the values to remain proportional to
the total number of opportunities (490 physicians accessible in the
region). As the decay behaviour decreases, more opportunities are
accessible for all zones. In the medium decay scenario \(f_2(c_{ij})\),
zone 1 sees a slight increase in values (relative to the highest decay
scenario) as the zone can accessible more opportunities relative to
increases seen in other zones. However, in the lowest decay scenario,
zone 1 sees a decrease, as it is outpaced by increases in other zones -
namely zone 2 (recall: zone 2 has the lowest number of opportunities,
hence the increases in opportunity gains is much higher in a low decay
scenario).

@@

Where the following \(\hat \kappa_i^T\) proportional allocation factor
is dimensionless:

\[
\hat \kappa_i^T = \frac{\sum_i O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

@@

\caption{\label{tbl-simple-example-total-population-accessibility}Simple
system: total opportunity constrained accessibility.}

@@

\begin{equation}\phantomsection\label{eq-opportunity-constrained-proportional-allocation-factor}{
\kappa^S_{i} = \frac{W_i^{(1)} f(c_{ij})}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}{
V^S_{i} = \kappa^S_{i} \sum_j   D_j
}\end{equation}

Soukhov et al. (2023) have shown that the role of \(\kappa^S_{i}\) is to
allocate opportunities \(D_j\) proportionally to the mass at each origin
\(i\) and the impedance between \(i\) and \(j\). As in the total
constrained opportunity case, \(\kappa^S_{i}\) is dimensionless and
\(V_i^S\) is in the units of opportunities \(D_j\). The singly
constrained accessibility measure in
Equation~\ref{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}
is called spatial availability by Soukhov et al. (2023), because it
represents the number of opportunities that can be reached \emph{and}
are available, in the sense that accessible opportunities have been
proportionally allocated based on relative demand, travel impedance and
the regional total number of opportunities, i.e., spatial competition
for them has been considered. These authors also show that the following
expression (accessibility per capita) is a constrained version of the
popular two-stage floating catchment area measure of Q. Shen (1998b) and
Luo and Wang (2003):

@@

\caption{\label{tbl-simple-example-attraction-constrained-accessibility}Simple
system: singly constrained opportunity accessibility.}

@@

Imposing the single proportional allocation factor \(\kappa^S_{i}\)
allows for the comparison of differences and ratios of the accessibility
values, like previously discussed in the total constrained accessible
opportunities case. The proportional allocation factor ensures that
resulting values are in units of \emph{physicians}, with the impedance
units already accounted for in the allocation process.

However, unlike the total constrained opportunity case, \(\kappa^S_{i}\)
reflects zonal competition based on the mass of the origin, i.e.,
population. Again, consider the highest decay scenario \(f_1(c_{ij})\).
Under this scenario, zone 1 no longer captures a medium amount of
physicians as in the total constrained opportunity case: it now captures
the fewest in the region i.e., 133.4687282 at zone 1, 166.7813387 at
zone 2, and 189.7499331 at zone 3. Why may zone 1 (healthcare cluster at
the edge of the city) capture 27\% of the physicians regionally while
this same zone captures 35\% in the total opportunity constrained case?

This difference is due to the single-opportunity factor
\(\kappa^S_{i}\)'s role in \(V_i^S\). The only inputs required in the
total-opportunity factor \(\kappa^T_i\) is the total number of
opportunities in the region \(D\) as well as the associated
opportunities at each \(j\) and travel impedance. Opportunities are
allocated to \(i\)s, regardless of the mass weights of origin. Whereas
the single opportunity constraint \(\kappa^S_{i}\) requires the
population at \(i\) as an input. In fact, \(\kappa^S_{i}\) is calculated
as the proportion of impedance-weighted population at an \(i\) to the
sum of impedance-weighted population for the entire region. Hence, since
zone 1 has the lowest population in the region, is in close proximity to
a more populated zone (zone 3, the `urban core'), and is not well
connected (in terms of travel impedance) to other zones with
opportunities, \(V_{1}^S\) values, or the number of physicians
accessible, is lowest.

@@

\caption{\label{tbl-simple-example-attraction-constrained-accessibility-per-capita}Simple
system: singly constrained opportunity accessibility per capita.}

@@

\begin{equation}\phantomsection\label{eq-production-constrained-accessibility-with-proportional-allocation-factor}{
M^S_{j} = \hat \kappa^S_{j} \sum_i  O_i
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-attraction-constrained-proportional-allocation-factor}{
\kappa^S _{j} = \frac{W_j^{(2)} f(c_{ij})}{\sum_i W_j^{(2)} f(c_{ij})}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-doubly-constrained-accessibility-w-paf}{
V^D_{i} = \kappa^D_{ij} \sum_j D_j
}\end{equation}

@@

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-f1cij-access-values}Doubly
constrained opportunity accessibility assuming the most decay travel
behaviour scenario in the modified simple system.}

@@

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-allfs-access-values-for-zone2}Doubly
constrained opportunity accessibility for all travel behaviour scenarios
for Zone 2 in the modified simple system.}

>>>>

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}
\usepackage{float}
\floatplacement{table}{H}
\floatplacement{figure}{H}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A family of accessibility measures derived from spatial interaction principles},
  pdfauthor={Anastasia Soukhov; Rafael H. M. Pereira; Christopher D. Higgins; Antonio Páez},
  pdfkeywords={accessibility; spatial interaction model; balancing
factors; competitive accessibility; constraints; singly constrained;
doubly constrained; total constrained},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{A family of accessibility measures derived from spatial
interaction principles}
\author{Anastasia Soukhov \and Rafael H. M. Pereira \and Christopher D.
Higgins \and Antonio Páez}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Transportation planning has long prioritized the efficiency of movement,
or mobility. However, the concept of accessibility represents a more
comprehensive evolution, shifting focus from mere movement to the
potential to reach (i.e., spatially interact) with desired destinations.
Despite growing recognition of accessibility-based planning approaches,
the concept remains fragmented, with inconsistent definitions and
unclear interpretations. This work's aim is to clarify and unify the
concept of accessibility by connecting it into spatial interaction
modeling. We demonstrate that widely used mobility and accessibility
models, such as gravity-based accessibility and spatial interaction
models, share common theoretical roots. From this foundation, this paper
offers three contributions: (A) we introduce a family of accessibility
measures within the principles of spatial interaction, and (B) formally
define four members of the family, namely the `unconstrained' measure
(i.e., Hansen-type accessibility), the `total constrained' measure
(i.e., a constrained version of the Hansen-type accessibility), the
`singly constrained' measure (i.e., related to the popular two step
floating catchment approach - 2SFCA), and the `doubly constrained'
measure representing realized interactions or `access', effectively
equal to the doubly constrained spatial interaction model; and (C) we
demonstrate the interpretability advantages of the family, as these
constrained accessibility measures yield values in units of the number
of potential ``opportunities for spatial interaction'' or ``population
for spatial interaction'' for each zone and zonal flow. The family of
accessibility measures proposed here clarifies the concept of
`potential' in accessibility, demonstrates theoretical and formulaic
linkages across popular accessibility and spatial interaction models,
and reintroduces measurement units into accessibility measures. By doing
so, we believe this family of measures can unlock a clearer, more
interpretable, and cohesive foundation for accessibility analysis.
\end{abstract}

@@

Understanding \(S_i\) as the inverse of Wilson's \(A_i\) does not
uncover any new meaning for \(S_i\) itself. Indeed, mathematically it is
true, \(A_i\)'s role in Wilson's general model
\(T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})\) is that of a balancing
factor \(k\) i.e., keeping units balanced and proportionality based on
constraints. Understanding \(S_i\) itself as a balancing factor is not
wholly helpful as Hansen and (Stewart before him) defined accessibility
as a partial sum of the demographic force \(F\) or the system-wide
population potential of opportunities for spatial interaction (i.e.,
Equation~\ref{eq-stewart-population-potential-sum} representing
\(V_i = \sum_j \frac{M_j}{d_{ij}}\) after losing \(G\)).

@@

Firstly, Table~\ref{tbl-small-system-land-use} summarises the population
(in units of 10,000s of people) and the opportunities (the number of
physicians) per zone. For reference, considering
Table~\ref{tbl-small-system-land-use}'s values, the
Provider-to-Population-Ratio (PPR) in this system is 24.5. The number of
physicians per 10,000 in Canada in 2022 was 24.97 (WHO 2025). Secondly,
to pair the zonal population and opportunity information, the assumed
cost of movement (in minutes of travel time) between origins and
destinations is as shown in Table~\ref{tbl-small-system-cost}.

@@

Further, we can see that, since \(D\) and \(W^{(2)}_j\) are both in
units of opportunities, the proportional allocation factor for the total
constrained opportunity case \(\kappa_{ij}^T\) is dimensionless:

\[
\kappa_{ij}^T = \frac{D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

\noindent and therefore \(V^T_i\) is now in the units of \(W^{(2)}_j\),
that is, the mass at the destination
(\(V^T_i = \sum_j \kappa_{ij}^T W^{(2)}_j\)). The role of
\(\kappa_{ij}^T\) in this reformulation of accessibility is to transform
between units and to adjust the number of opportunities accessible from
\(i\) so they represent a proportion of the total number of
opportunities in the region. \(\kappa_{ij}^T\) then assigns
opportunities in proportion to the impedance between \(i\) and \(j\).
This is why we refer to \(\kappa_{ij}^T\) as a proportional allocation
factor. On the other hand, the proportionality constant \(K^T\) balances
the units of \(V^0_i\), the Hansen-type accessibility values, and is an
alternative expression of the total constrained accessibility measure.

@@

Using the calculated proportionality constants for all zones and
multiplying them by the unconstrained accessibility value \(V^0_i\), the
total opportunity constrained accessibility values for all zones and
different travel behaviour scenarios is presented in
Table~\ref{tbl-simple-example-total-opportunity-accessibility}.
\(\kappa_{ij}^T\) for each zone is not shown, but can be understood to
be the unitless proportion of opportunities (of the total opportunities)
allocated to each zone based on travel impedance.

@@

\caption{\label{tbl-simple-example-total-opportunity-accessibility}Simple
system: total constrained accessible opportunities.}

@@

In contrast to unconstrained accessibility \(V^0_i\), imposing a
constraint -in this case a total opportunity constraint for this
variant- allows for the comparison of differences and ratios between
regions and across different travel behaviour scenarios as well. Each
value is effectively in units of physicians, with the impedance units
already accounted for by \(\kappa_{ij}^T\).

@@

One can also directly compare values at a specific zone due to the
consistent units. For instance, zone 1 remains intermediate in capturing
accessible physicians relative to zones 2 and 3 across scenarios,
similar to the unconstrained case. However, the difference between
travel behaviour scenarios differ in direction. Specifically, Zone 1
captures 0.6067946 more and 7.9850478 fewer opportunities than the
lesser decay scenarios \(f_2(c_{ij})\) and \(f_3(c_{ij})\) respectively.
Why? \(\kappa_{ij}^T\) ensures proportional allocation for each travel
behaviour scenario. Meaning, while the unconstrained accessibility
increases, \(\kappa_{ij}^T\) adjusts the values to remain proportional
to the total number of opportunities (490 physicians accessible in the
region). As the decay behaviour decreases, more opportunities are
accessible for all zones. In the medium decay scenario \(f_2(c_{ij})\),
zone 1 sees a slight increase in values (relative to the highest decay
scenario) as the zone can accessible more opportunities relative to
increases seen in other zones. However, in the lowest decay scenario,
zone 1 sees a decrease, as it is outpaced by increases in other zones -
namely zone 2 (recall: zone 2 has the lowest number of opportunities,
hence the increases in opportunity gains is much higher in a low decay
scenario).

@@

Where the following \(\hat \kappa_{ji}^T\) proportional allocation
factor is dimensionless:

\[
\hat \kappa_{ji}^T = \frac{O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

@@

\caption{\label{tbl-simple-example-total-population-accessibility}Simple
system: total constrained accessible population.}

@@

\begin{equation}\phantomsection\label{eq-opportunity-constrained-proportional-allocation-factor}{
\kappa^S_{ij} = \frac{W_i^{(1)} f(c_{ij})}{\sum_i W_i^{(1)} f(c_{ij})}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}{
V^S_{i} = \sum_j \kappa^S_{ij} D_j
}\end{equation}

Soukhov et al. (2023) have shown that the role of \(\kappa^S_{ij}\) is
to allocate opportunities \(D_j\) proportionally to the mass at each
origin \(i\) and the impedance between \(i\) and \(j\). As in the total
constrained opportunity case, \(\kappa^S_{ij}\) is dimensionless and
\(V_i^S\) is in the units of opportunities \(D_j\). The singly
constrained accessibility measure in
Equation~\ref{eq-attraction-constrained-accessibility-with-proportional-allocation-factor}
is called spatial availability by Soukhov et al. (2023), because it
represents the number of opportunities that can be reached \emph{and}
are available, in the sense that accessible opportunities have been
proportionally allocated based on relative demand, travel impedance and
the regional total number of opportunities, i.e., spatial competition
for them has been considered. These authors also show that the following
expression (accessibility per capita) is a constrained version of the
popular two-stage floating catchment area measure of Q. Shen (1998b) and
Luo and Wang (2003):

@@

\caption{\label{tbl-simple-example-attraction-constrained-accessibility}Simple
system: singly constrained accessible opportunities.}

@@

Imposing the single proportional allocation factor \(\kappa^S_{ij}\)
allows for the comparison of differences and ratios of the accessibility
values, like previously discussed in the total constrained accessible
opportunities case. The proportional allocation factor ensures that
resulting values are in units of \emph{physicians}, with the impedance
units already accounted for in the allocation process.

However, unlike the total constrained opportunity case,
\(\kappa^S_{ij}\) reflects zonal competition based on the mass of the
origin, i.e., population. Again, consider the highest decay scenario
\(f_1(c_{ij})\). Under this scenario, zone 1 no longer captures a medium
amount of physicians as in the total constrained opportunity case: it
now captures the fewest in the region i.e., 133.4687282 at zone 1,
166.7813387 at zone 2, and 189.7499331 at zone 3. Why may zone 1
(healthcare cluster at the edge of the city) capture 27\% of the
physicians regionally while this same zone captures 35\% in the total
opportunity constrained case?

This difference is due to the single-opportunity factor
\(\kappa^S_{ij}\)'s role in \(V_i^S\). The only inputs required in the
total-opportunity factor \(\kappa^T_{ij}\) is the total number of
opportunities in the region \(D\) as well as the associated
opportunities at each \(j\) and travel impedance. Opportunities are
allocated to \(i\)s, regardless of the mass weights of origin. Whereas
the single opportunity constraint \(\kappa^S_{ij}\) requires the
population at \(i\) as an input. In fact, \(\kappa^S_{ij}\) is
calculated as the proportion of impedance-weighted population at an
\(i\) to the sum of impedance-weighted population for the entire region.
Hence, since zone 1 has the lowest population in the region, is in close
proximity to a more populated zone (zone 3, the `urban core'), and is
not well connected (in terms of travel impedance) to other zones with
opportunities, \(V_{1}^S\) values, or the number of physicians
accessible, is lowest.

@@

\caption{\label{tbl-simple-example-attraction-constrained-accessibility-per-capita}Simple
system: singly constrained accessible opportunities per capita.}

@@

\begin{equation}\phantomsection\label{eq-production-constrained-accessibility-with-proportional-allocation-factor}{
M^S_{j} = \hat \kappa^S_{ji} \sum_i  O_i
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-attraction-constrained-proportional-allocation-factor}{
\hat \kappa^S_{ji} = \frac{W_j^{(2)} f(c_{ij})}{\sum_i W_j^{(2)} f(c_{ij})}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-doubly-constrained-accessibility-w-paf}{
V^D_{i} = \sum_j \kappa^D_{ij} D_j
}\end{equation}

@@

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-f1cij-access-values}Doubly
constrained accessible opportunities assuming highest travel decay in
the modified simple system.}

@@

\caption{\label{tbl-adjusted-small-system-land-use-doubly-constrained-case-allfs-access-values-for-zone2}Doubly
constrained accessible opportunities at Zone 2 for all travel decay
groups in the modified simple system.}

### 2025-05-23 paper/summary-family-access-measures-table.tex
Cleaned up a few small equation errors - and posted preprint link to repo page.

{\tiny
\begin{longtable}{|p{2.5cm}|p{2.5cm}|p{2.5cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Name of Case and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endfirsthead

\hline
\textbf{Name of Case and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endhead

Unconstrained Accessible Opportunities ($V_i^0$) and Unconstrained Accessibile Population ($M_j^0$)
& No constraints; marginals not equal to any regional or zonal knowns.
& None
& $V_i^0 = \sum_j O_j f(c_{ij})$;

$M_j^0 = \sum_i P_i f(c_{ij})$

@@

& Allocates the total marginal as opportunities based on $\kappa_i^T = \sum_j K^T f(c_{ij})$ and as population based on  $\hat{\kappa}_j^T = \sum_i \hat K^T f(c_{ij})$

& $V^T_i = \kappa_i^T \sum_j W^{(2)}_j$;

$M_j^T = \hat{\kappa}_j^T \sum_i W^{(1)}_i$

@@

& Allocates the single opportunities marginal proportionally based on $\kappa^S_i = W_i^{(1)} B_j$ in the case of $V_i^S$ and the single population marginal proroportionally based on $\hat \kappa^S_j = W_j^{(2)} A_i$ in the case of $M_j^S$

& $V^S_i = \kappa^S_i \sum_j D_j$;

$M_j^S = \hat \kappa^S_j \sum_i P_i$

@@

Doubly Constrained Access ($V_{ij}^D$ or $M_{ij}^D$)
& Values reflect both single marginals simutaneously, maintained via $A_i$ and $B_j$.
& —
& $V_{ij}^D = A_i B_j P_i O_j f(c_{ij})$
& The spatial interactions between population and opportunities (i.e., access). \\
\hline

>>>>

{\tiny
\begin{longtable}{|p{2.5cm}|p{2.5cm}|p{2.5cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Name of Member and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endfirsthead

\hline
\textbf{Name of member and Variant} & \textbf{Constraint Explanation and Balancing Factor} & \textbf{Proportional Allocation Factor} & \textbf{Measure Equation} & \textbf{Interpretation} \\
\hline
\endhead

Unconstrained Accessible Opportunities ($V_i^0$) and Unconstrained Accessibile Population ($M_j^0$)
& No constraints; marginals not equal to any regional or zonal knowns.
& None
& $V_i^0 = \sum_j D_j f(c_{ij})$;

$M_j^0 = \sum_i O_i f(c_{ij})$

@@

& Allocates the total marginal as opportunities based on $\kappa_{ij}^T = K^T f(c_{ij})$ and as population based on  $\hat{\kappa}_{ji}^T = \hat K^T f(c_{ij})$

& $V^T_i =  \sum_j \kappa_{ij}^T W^{(2)}_j$;

$M_j^T =  \sum_i \hat{\kappa}_{ji}^T W^{(1)}_i$

@@

& Allocates the single opportunities marginal proportionally based on $\kappa^S_{ij} =  B_j W_i^{(1)} f(c_{ij})$ in the case of $V_i^S$ and the single population marginal proroportionally based on $\hat \kappa^S_{ji} = A_i W_j^{(2)} f(c_{ij})$ in the case of $M_j^S$.

& $V^S_i = \sum_j \kappa^S_{ij} D_j$;

$M_j^S =  \sum_i \hat \kappa^S_{ji} O_i$

@@

Doubly Constrained Access ($V_{ij}^D$ or $M_{ij}^D$)
& Values reflect both single marginals simutaneously, maintained via $A_i$ and $B_j$.
& —
& $V_{ij}^D = A_i B_j O_i D_j f(c_{ij})$
& The spatial interactions between population and opportunities (i.e., access). \\
\hline
